Phenology and Mycorrhizae Manuscript

### 2022-12-06 DataExploratoryAnalyses.Rmd
More exploratory analyses

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
```

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
library(MuMIn)
library(ggplot2)
library(mgcv)
library(ggridges)

allPrecipData <- read.csv(paste0(dataPath,"NEON_PrecipitationDataDailydata_gapFilled.csv"))
allPrecipData$year <- lubridate::year(allPrecipData$date)
allPrecipData <- allPrecipData %>% group_by(year,siteID) %>% mutate(cumPrecip=cumsum(precipBulk)) #Calculates the cummulative sum of precip for each site-year for each date

@@

#For each site-phenophase combination: how many years are individuals re-measured?
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=10}
for(p in seq_along(NEON_phenophase_names)){
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','year')]
  subDat <- subDat %>% unique()

@@

#For each PFT-phenophase combination: Does the Standard Deviation in DOYs and lengths between phenophases differ between EcM and AM associated plants acrross all sites and within each site?

@@

#Lengths Between Phenophases:

@@

#For each PFT-phenophase combination: Does the DOY of the timing and the lengths between phenophases differ between EcM and AM assocated plants across all sites and within each site?

@@

>>>>

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
library(MuMIn)
library(ggplot2)
library(mgcv)
library(ggridges)
```

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
allPrecipData <- read.csv(paste0(dataPath,"NEON_PrecipitationDataDailydata_gapFilled.csv"))
allPrecipData$year <- lubridate::year(allPrecipData$date)
allPrecipData <- allPrecipData %>% group_by(year,siteID) %>% mutate(cumPrecip=cumsum(precipBulk)) #Calculates the cummulative sum of precip for each site-year for each date

@@

#For each phenophase, what growth forms (PFT) are measured?
```{r,echo=FALSE,message=FALSE,warning=FALSE}
allDat <- matrix(nrow=0,ncol=6)
for(p in seq_along(NEON_phenophase_names)){
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat[,c('siteID','individualID','growthForm','decimalLatitude','year','phenophaseName')]
  subDat <- subDat %>% unique() %>% filter(!is.na(growthForm))
  allDat <- rbind(allDat,subDat)
}
  allDat <- allDat %>% filter(growthForm %in% c("Deciduous broadleaf","Drought deciduous broadleaf","Evergreen broadleaf","Evergreen conifer","Forb","Graminoid","Semi-evergreen broadleaf"))

  growthFormTotal <- allDat %>% group_by(growthForm) %>% summarise(growthFormTotal=n())
  allDat <- left_join(allDat,growthFormTotal,by="growthForm")

  allDat2 <- allDat %>% group_by(phenophaseName,growthForm,growthFormTotal) %>% summarise(n=(n()))
  # ggplot(allDat2,aes(n,phenophaseName,fill=growthForm)) +
  #           geom_bar(position="dodge",stat="identity")+
  #           scale_fill_brewer(palette="Accent")+
  #           theme_classic()+
  #           theme(axis.text.x = element_text(angle=90))+
  #           xlab("Number Of Total Measurements (Ratio of Total Per Growth Form")+
  #           ylab("Phenophase")+
  #           ggtitle(NEON_phenophase_names[p])

    ggplot(allDat2,aes((n/growthFormTotal),growthForm,fill=phenophaseName)) +
            geom_bar(position="dodge",stat="identity")+
            scale_fill_brewer(palette="Set3")+
            theme_classic()+
            theme(axis.text.x = element_text(angle=90))+
            xlab("Number Of Total Measurements (Ratio of Total Per Growth Form)")+
            ylab("Growth Form")+
            ggtitle("Which Phenophases are Monitored for with Growth Forms")

#For each site-phenophase combination: how many years are individuals re-measured by mycorrhizae type?
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=10}
for(p in seq_along(NEON_phenophase_names)){
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','year')]
  subDat <- subDat %>% unique()

@@

#For each PFT-phenophase combination: Does the timing and Standard Deviation in DOYs differ between phenophases differ between EcM and AM associated plants acrross all sites and within each site?

@@

@@

#For each PFT-phenophase combination: Does the length and Standard Deviation differ between phenophases differ between EcM and AM associated plants acrross all sites and within each site?

##There are various definitions of growing season length depending on PFT:
*Semi-evergreen broadleaf: budburst-colored; leaves (varying intensities); Young-leaves-colored
*Graminoid: leaves (varying intensities)
*Forb: leaves (varying intensities); Young-leaves-colored
*Evergreen conifer: No lengths
*Evergreen broadleaf: No lengths
*Drought deciduous broadleaf: leaves (varying intensities); young leaves-colored leaves
*Deciduous broadleaf: leaves (varying intensities); budburst-colored

##Different definitions of growing season length:
*Budburst-colored: semi-evergreen broadleaf; deciduous broadleaf
*Leaf presence: semi-evergreen broadleaf; graminoid; forb; drought deciduous broadleaf; deciduous broadleaf
*Young-leaves-colored:semi-evergreen broadleaf; forb; drought-deciduous broadleaf

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE,fig.width=10,fig.height=20}
for(initialP in c(1,2)){
  p=initialP
  budBurstDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  budBurstDat <- budBurstDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])

  budBurstDat <- budBurstDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','dayOfYear','growthForm','year','date')] %>% mutate(winterYear=as.numeric(year),year=as.numeric(year))

  p=3
  coloredLeavesDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  coloredLeavesDat <- coloredLeavesDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])

  coloredLeavesDat <- coloredLeavesDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','dayOfYear','growthForm','year','date')]%>% mutate(winterYear=(year+1),year=as.numeric(year))

  winterLengthDat <- inner_join(coloredLeavesDat,budBurstDat,
                            by=c("winterYear","individualID",'siteID',"Mycorrhizal.type","decimalLatitude",
                                     "growthForm")) %>% filter(growthForm %in% c("Deciduous broadleaf","Drought deciduous broadleaf","Evergreen broadleaf","Evergreen conifer","Forb","Graminoid","Semi-evergreen broadleaf"))%>% 
    mutate(winterLength=as.Date(date.y)-as.Date(date.x)) %>%
    group_by(individualID) %>% mutate(winter_sd=sd(winterLength,na.rm=TRUE)) %>%
    ungroup() %>%
    dplyr::select(individualID,winter_sd,siteID,Mycorrhizal.type,decimalLatitude,growthForm) %>% unique()

  summerLengthDat <- inner_join(coloredLeavesDat,budBurstDat,
                                by=c("year","individualID",'siteID',"Mycorrhizal.type",
                                     "decimalLatitude","growthForm")) %>% filter(growthForm %in% c("Deciduous broadleaf","Drought deciduous broadleaf","Evergreen broadleaf","Evergreen conifer","Forb","Graminoid","Semi-evergreen broadleaf")) %>%
    mutate(summerLength=as.Date(date.x)-as.Date(date.y)) %>%
    group_by(individualID) %>% mutate(summer_sd=sd(summerLength,na.rm=TRUE))%>%
    ungroup() %>%
    dplyr::select(individualID,summer_sd,siteID,Mycorrhizal.type,decimalLatitude,growthForm) %>% unique()

  print(ggplot(winterLengthDat,
         aes(winter_sd,fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
    geom_density_ridges2(scale=1)+
    theme_bw()+
    theme(axis.text.x = element_text(angle=90))+
    xlab("SD")+
    ylab("Site ID (Ordered by Latitude)")+
    facet_wrap(~growthForm)+
    ggtitle(paste("SD in Winter Length Per Individual: Colored-",NEON_phenophase_names[initialP]))+
    scale_x_continuous(limits=c(0,100)))
}
# #Based on Leaf Numbers
# leavesDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
# initialGrowth <- leavesDat %>% filter(phenophaseIntensity=="5-")

@@

```{r}
p=5

### 2022-12-06 manuscriptCode_phenologyAndMycorrhizae.Rmd
More exploratory analyses

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
combineNEONdata_PhenoObs(saveType=="bySite") #Saved in individual files by site because 1 file was too large 
combineNEONdata_PhenoObs(saveType=="firstOfPhenophase")
```

>>>>

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
combineNEONdata_PhenoObs(saveType=="bySite") #Saved in individual files by site because 1 file was too large 
combineNEONdata_PhenoObs(saveType=="firstOfPhenophase")
combineNEONdata_PhenoObs(saveType = "allOfPhenophase")
```
